# Evaluation

##### Each ATE project is required to have an evaluation component to assess its quality and effectiveness. Evaluation of ATE and other NSF-funded projects is intended to serve two distinct purposes: (1) Produce information that can be used to improve a project as it is being implemented and (2) Determine and document a project’s achievements [@frechtling2010user].

##### ATE PIs were asked about their evaluators and interactions with them, as well as their projects’ use and dissemination of evaluation results.

## Evaluation

```{r 11.1_prep}
ATE_evaluation_summary <- 
  ATE_data %>% 
  process_summary(
    yes_no_planning_mapping,
    'Q10.1. Did your project have an evaluator in 2020?'
  )

evaluator_type_mapping = data.frame(
  number = 1:3,
  evaluator_type = c('External evaluator only', 'Internal evaluator','Both internal and external evaluators')
)

ATE_evaluator_type_mapping <- ATE_data %>% 
  process_summary(
    evaluator_type_mapping,
    'Q10.2. Which type of evaluator(s) did your project have in 2020?'
  )

evaluator_interaction_frequency_mapping <- data.frame(
  number = 1:5,
  evaluator_interaction_frequency = c(
    'Rarely (annually or semiannually)', 
    'Infrequently (not every month but at least quarterly)',
    'Occasionally (more often than quarterly and as much as monthly)',
    'Often (more often than monthly and as much as biweekly)',
    'Continually (very nearly weekly, weekly, or more often)'
  )
)

ATE_evaluator_interaction_frequency_summary <-
  ATE_data %>% 
  process_summary(
    evaluator_interaction_frequency_mapping,
    "Q10.3. How frequently did your external evaluator interact with your project's staff (e.g., by email, teleconference, face-to-face) in 2020?"
  )

evaluation_report_type_mapping <- data.frame(
  number = 1:4,
  evaluator_interaction_frequency = c(
    'Written', 
    'Oral',
    'Both oral and written',
    'None'
  )
)

ATE_evaluation_report_type_summary <-
  ATE_data %>% 
  process_summary(
    evaluation_report_type_mapping,
    'Q10.4. What type of report did you receive from your evaluator in 2020?'
)


evaluation_caused_goals_changes_total <-
  ATE_data %>%
  select(starts_with('Q10.8. '), -ends_with('Text')) %>%
  filter(rowSums(. == 1, na.rm = TRUE) > 0) %>%
  nrow()

ATE_evaluation_shared_results_to_summary <- ATE_data %>% 
  process_summary_2(starting_with = 'Q10.8. ', names_to = 'shared_results_to', summary = 'count_ones') %>% 
  mutate(pct = count_ones/evaluation_caused_goals_changes_total * 100)
```

```{r 11.1_inline}
# paragraph 1
projects_engaged_evaluator_pct_en <- 
  ATE_evaluation_summary %>%
  filter(yes_no_planning == 'Yes') %>%
  pull(pct) %>%
  round_and_english()

# paragraph 2
projects_engaged_evaluator_en <- 
  ATE_evaluation_summary %>%
  filter(yes_no_planning == 'Yes') %>%
  pull(n) %>%
  round_and_english()
no_projects_engaged_evaluator <- 
  ATE_evaluation_summary %>% 
  filter(yes_no_planning == 'No') %>%
  pull(n) %>%
  round2()
no_projects_engaged_evaluator_first_year_en <- 
  ATE_data %>%
  filter(`Q1.2. Was 2020 the first year of your current grant?` == 1, 
         `Q10.1. Did your project have an evaluator in 2020?` == 5) %>%
  nrow()
projects_engaged_evaluator <- 
  ATE_evaluation_summary %>%
  filter(yes_no_planning == 'Yes') %>%
  pull(n) %>%
  round2()
external_evaluator_pct <- 
  ATE_evaluator_type_mapping %>%
  filter(evaluator_type == 'External evaluator only') %>%
  pull(pct) %>%
  round2()
internal_and_external_evaluator_pct <- 
  ATE_evaluator_type_mapping %>%
  filter(evaluator_type == 'Both internal and external evaluators') %>%
  pull(pct) %>%
  round2()
internal_evaluator_pct <- 
  ATE_evaluator_type_mapping %>%
  filter(evaluator_type == 'Internal evaluator') %>%
  pull(pct) %>%
  {2} %>%
  round2()

# paragraph 3
continually_or_often <- 
  ATE_evaluator_interaction_frequency_summary %>%
  filter(str_detect(evaluator_interaction_frequency, "^Continually") | 
         str_detect(evaluator_interaction_frequency, "^Often")) %>%
  pull(pct) %>%
  sum() %>% 
  round_and_english()
occasionally <- 
  ATE_evaluator_interaction_frequency_summary %>%
  filter(str_detect(evaluator_interaction_frequency, "^Occasionally")) %>%
  pull(pct) %>%
  sum() %>% 
  round2()
infrequently <- 
  ATE_evaluator_interaction_frequency_summary %>%
  filter(str_detect(evaluator_interaction_frequency, "^Infrequently")) %>%
  pull(pct) %>%
  sum() %>% 
  round2()

# figure 22
total_evaluation_report_type <- sum(ATE_evaluation_report_type_summary$n)

# paragraph 4
total_received_evaluation_report <- ATE_evaluation_report_type_summary %>% 
  filter(evaluator_interaction_frequency != 'None') %>% 
  pull(n) %>% 
  sum()
# evaluation_caused_implementation_changes_pct <-
#   ATE_data %>%
#   filter(`Q10.4. What type of report did you receive from your evaluator in 2020?` != 'None') %>%
#   select(starts_with('Q10.6. '), -ends_with('Text')) %>%
#   filter(rowSums(!is.na(.)) != 0) %>%
#   nrow() / total_received_evaluation_report * 100
# evaluation_caused_goals_changes_pct <-
#   ATE_data %>%
#   filter(`Q10.4. What type of report did you receive from your evaluator in 2020?` != 'None') %>%
#   select(starts_with('Q10.7. '), -ends_with('Text')) %>%
#   filter(rowSums(!is.na(.)) > 0) %>%
#   nrow() / total_received_evaluation_report * 100
```

#### `r projects_engaged_evaluator_pct_en` percent of ATE projects engaged an evaluator.

`r projects_engaged_evaluator_en` ATE projects had evaluators in 2023. Of the `r no_projects_engaged_evaluator` PIs who said they did not have evaluators, 12 were in their first year of funding. **Of the `r projects_engaged_evaluator` projects with evaluators, `r external_evaluator_pct`% reported having an external evaluator, with `r internal_and_external_evaluator_pct`% having both an internal and external evaluator and `r internal_evaluator_pct`% having only an internal evaluator.**

`r continually_or_often` percent of PIs reported that they interacted with their evaluators continually (at least once a week) or often (two or three times a month), while `r occasionally`% interacted with their evaluators occasionally (more often than quarterly) and `r infrequently`% did so infrequently or rarely (once a quarter or less).

<!-- manual update -->
##### More than half of ATE projects received `r colorize('both oral and written', color1)` evaluation reports.
<!-- manual update -->

```{r fig11a1, fig.cap = paste0("Types of evaluation report received by ATE projects (n=", total_evaluation_report_type, ")"), fig.height=2}
set_color_palette(ATE_evaluation_report_type_summary)
create_bar_chart_single(ATE_evaluation_report_type_summary, pct, evaluator_interaction_frequency,  elevate_labels = FALSE)

```

<!-- manual update -->
**Of the `r total_received_evaluation_report` PIs who received evaluation reports, 53% indicated their project’s evaluation caused them to make a change in implementing their project,** and 46% indicated that the evaluation caused them to make a change in their project’s goals, objectives, or target audience.
<!-- manual update -->

\newpage

<!-- manual update -->
##### Most projects shared their evaluation results with `r colorize('NSF program officers', color1)`, their `r colorize('faculty or staff at their host institution', color2)`, and `r colorize('executive administrators', color3)`.
<!-- manual update -->

```{r fig11a2, fig.cap = paste0("Percentage of projects that shared their evaluation results with various audiences (n=", evaluation_caused_goals_changes_total, ")"), fig.height = length(ATE_evaluation_shared_results_to_summary$shared_results_to) * 0.75}
set_color_palette(ATE_evaluation_shared_results_to_summary)
create_bar_chart(ATE_evaluation_shared_results_to_summary, pct, shared_results_to)
```

