# ATE Survey 2024 Highlights

```{r 12.1_prep}
ATE_students_served_by_projects_summary <- tibble(
  activities = c("Attended academic programs", "Completed courses",
                             "Participated in a transition program", "Received business and entrepreneurial skills development",
                             "Participated in student competitions", "Received mentoring",
                             "Engaged in workplace-based learning"),  
  n = as.integer(str_replace_all(c(total_students_attended, total_students_completed_course,
                                   total_students_transition_program, total_students_business_skills,
                                   total_students_competition, total_students_mentoring_coaching,
                                   total_students_wbl), ",", ""))  # Convert to integer
) %>% 
  mutate(activities = fct_reorder(activities, n, .desc = TRUE))

ATE_professional_development_educators_summary <- ATE_data %>%
  select(
    starts_with('Q4.2.') & (ends_with('High School Teachers') | 
                            ends_with('2-Year College Faculty') | 
                            ends_with('4-Year College Faculty'))
  ) %>% 
  pivot_longer(
    cols = everything(),
    names_to = c(".value"),
    names_pattern = "(High School Teachers|2\\-Year College Faculty|4\\-Year College Faculty)$"
  ) %>% 
  summarize(across(everything(), ~ sum(.x, na.rm = TRUE))) %>% 
  pivot_longer(cols = everything(), names_to = 'educators', values_to = 'n') %>% 
  mutate(
    educators = fct_reorder(educators, n, .desc = TRUE),
    n = round(n, -1),  # Round n to the nearest ten
    pct = n/sum(n) * 100
    ) %>% 
  arrange(desc(n))

ATE_articulation_detail_summary_updated <- ATE_articulation_detail_summary %>%
  filter(str_detect(detail, 'Articulation')) %>%
  group_by(detail) %>%
  mutate(pct = (n / sum(n)) * 100,
         detail = str_replace(detail, "Articulation agreements in place in 2020", "Agreements Maintained"),
         detail = str_replace(detail, "Articulation agreements developed", "Agreements Developed"),
         detail = fct_inorder(detail),
         education = str_replace(education, "High school to Two-year college", "High school to  \n Two-year college"),
         education = str_replace(education, "Two-year college to Four-year college", "Two-year college to  \n Four-year college"))

collaboration_business <- ATE_collaboration_organizations_summary %>%
  filter(organization == 'Business/industry') %>%
  add_row(organization = '', pct = 100 - .$pct) %>%
  mutate(organization = fct_inorder(organization))

collaboration_college <- ATE_collaboration_organizations_summary %>%
  filter(organization == '2- or 4-year colleges') %>%
  add_row(organization = '', pct = 100 - .$pct) %>%
  mutate(organization = fct_inorder(organization))



total_primary_purpose_coordination <- 
  ATE_coordination_network_summary %>%
  filter(yes_no_planning == 'Yes') %>%
  pull(n) 
```


```{r 12.1_inline}
# 130 DEGREE PROGRAMS AND 370 COURSES
conferences = ATE_grant_type_summary %>% 
  filter(str_detect(grant_type, 'Conference')) %>% 
  pull(n)
total_projects_on_degree_program_and_courses <- ATE_data %>%
  select(starts_with('Q2.2') & contains('Degree'),
         starts_with('Q2.10.') & ends_with('Audience')) %>% 
  rowwise() %>%  # Treat each row independently
  filter(any(!is.na(c_across(everything())))) %>%  # Filter rows with at least one non-NA value
  nrow()

# 8,630 EDUCATORS
two_year_college_faculty <- 
  ATE_professional_development_educators_summary %>%
  filter(str_detect(educators, '2-Year')) %>%
  pull(n) %>%
  round_to_ten()
four_year_college_faculty <- 
  ATE_professional_development_educators_summary %>%
  filter(str_detect(educators, '4-Year')) %>%
  pull(n) %>%
  round_to_ten()

# 39,500+ STUDENTS
total_students_all_activities <- ATE_students_served_by_projects_summary %>% 
  pull(n) %>% 
  sum() %>% 
  round_to_ten()

# 630 ARTICULATION AGREEMENTS
total_articulation_agreements_developed_maintained <- ATE_articulation_detail_summary %>% 
  filter(str_detect(detail, 'Articulation')) %>% 
  pull(n) %>% 
  sum() %>% 
  round_to_ten()

students_matriculated <- ATE_articulation_detail_summary %>% 
  filter(str_detect(detail, 'transferred')) %>% 
  pull(n) %>% 
  sum() %>% 
  round_to_ten()


total_collaborators <- ATE_collaboration_organizations_summary %>% 
  pull(sum) %>% 
  sum() %>% 
  round_to_ten()
```

<!-- manual update -->
This summary of activities and achievements of the Advanced Technology Education (ATE) program is based on the 2024 ATE Survey. Principal investigators for 97% of ATE grants (n= `r total_ATE_grants`, out of a total of 322 active ATE grants) completed the survey. Respondents represented grants supporting `r projects` projects, `r center` centers, `r conferences` conferences, and `r targeted_research` targeted research projects. 
<!-- manual update -->

\newpage

##### `r colorize(paste(total_programs, "degree programs and", total_developed_courses, "courses were developed by", total_projects_on_degree_program_and_courses, "ATE projects."), color7)`

<!-- manual update -->
Most (`r certificate_pct`%) of all academic degree programs developed were certificate programs, and a majority (`r developed_courses_two_year_college_audience`%) of courses developed were at the associate degree level. 
<!-- manual update -->



```{r, fig.height = 2}
icons_list <- c("&#xf0a3", "&#xf70e")
text_list <- c(paste0(certificate, " certificate programs served<br>", certificate_students, " students"),
               paste0(associate, " associate degree programs served<br>", associate_students, " students"))
colors_list <- c(color1,color2)

df <- create_callout_df(icons = icons_list, text = text_list, colors = colors_list)

create_callout_chart(df)
```

##### `r colorize(paste(total_educators_participated, "educators participated in", total_activities, "professional development activities."), color7)`

`r two_year_college_faculty` educators who received professional development were located at two-year colleges followed by `r four_year_college_faculty` educators at four-year colleges. 
<!-- manual update -->
<!-- *over* `r two_year_college_faculty` educators-->


```{r fig.height= 2 + 0.1 * nlevels(ATE_professional_development_educators_summary$educators)}
ATE_professional_development_educators_summary_filtered <- ATE_professional_development_educators_summary %>%
  filter(!(educators %in% c("2-Year College Faculty", "4-Year College Faculty", "High School Teachers")))
set_color_palette(ATE_professional_development_educators_summary)
create_bar_chart_single(ATE_professional_development_educators_summary, n, educators, pct = FALSE)
```

\newpage

##### `r colorize(paste(total_students_all_activities, "students were served by ATE projects."), color7)`

ATE projects served over `r total_students_all_activities` students through a variety of activities in 2020.* ATE projects served the most students through developing or modifying academic programs, followed by courses. 
<!-- manual update -->
<!-- `r total_students_all_activities`*+* -->
<!-- *over* `r total_students_all_activities` students -->

```{r, fig.height = length(ATE_students_served_by_projects_summary$activities) * 0.75}
set_color_palette(ATE_students_served_by_projects_summary)
create_bar_chart(ATE_students_served_by_projects_summary, n, activities, pct = FALSE)
```

* Due to the structure of the survey questions, the total number of students served may not represent a count of unique students.

\newpage

##### `r colorize(paste(total_educational_material_development, "educational materials were developed by", total_educational_material_development_projects, "ATE projects."), color7)`

Educational materials developed included assessment activities, modules or instructional units, lessons, lab experiments, curricula, case studies, instructor guides, and textbooks.

```{r, fig.height = 3}
icons_list <- c("&#xf15c", "&#xf303", "&#xf51c")
text_list <- c(paste0(assessment_activities, " Assessment activities or tests"),
               paste0(module_instructional_unit, " Modules or instructional units"),
               paste0(lesson_plan, " Lessons or lesson plans"))
colors_list <- c(color1, color2, color3)

df <- create_callout_df(icons = icons_list, text = text_list, colors = colors_list)


create_callout_chart(df)
```


##### `r colorize(paste(total_articulation_agreements_developed_maintained, "articulation agreements were developed or maintained by", total_articulation, "ATE projects."), color7)`

`r students_matriculated` students matriculated to a higher-level education institution with the aid of an ATE-supported articulation agreement.

```{r, fig.height=nlevels(ATE_articulation_detail_summary_updated$detail) * 1.6, fig.width = 8.5}
set_color_palette(ATE_articulation_detail_summary_updated)
create_bar_chart_proportional(ATE_articulation_detail_summary_updated, pct, detail, education)
```

\newpage

##### `r colorize(paste(total_collaborators, "collaborators were engaged by ATE projects."), color7)`

ATE projects most frequently collaborated with business and industry groups and other colleges and universities.

```{r, fig.width = 8, fig.height = 3}
p1 <- collaboration_business %>%
  ggplot(aes(fill = organization, values = round(pct))) +
  geom_waffle(size = 1, colour = "white", flip = TRUE) +
  scale_fill_manual(guide = "none", values = c('#783a6b','#e7e6e6'))+
  scale_x_continuous(expand = expansion(mult = c(0, .9)))+
  annotate("text", label = paste0(round(collaboration_business$pct[1]), "% \ncollaborated with \nbusiness and \nindustry groups"),
           x = 11, y = 10,
           color = "#783a6b", fontface = 'bold', size = 5, hjust = 0, vjust = 1) +
  coord_equal() +
  theme_void()

p2 <- collaboration_college %>%
  ggplot(aes(fill = organization, values = round(pct))) +
  geom_waffle(size = 1, colour = "white", flip = TRUE) +
  scale_fill_manual(guide = "none", values = c(color1,'#e7e6e6'))+
  scale_x_continuous(expand = expansion(mult = c(0, .9)))+
  annotate("text", label = paste0(round(collaboration_college$pct[1]), "% \ncollaborated with \nother two- and \nfour-year colleges"),
           x = 11, y = 10,
           color = color1, fontface = 'bold', size = 5, hjust = 0, vjust = 1) +
  coord_equal() +
  theme_void()

p1 + p2
```




##### `r colorize("Other activities were conducted by ATE projects in 2023 in support of advanced technological education.", color7)`
<!-- MANUAL UPDATE -->

ATE projects engaged in a wide range of activities. More information about those listed below and others can be found in the full report.


```{r, fig.height = 4}

icons_list <- c("&#xe068", "&#xf518", "&#xe4da", "&#xf1ea")
text_list <- c(paste0(total_primary_purpose_coordination, " coordination networks operated"),
               paste0(total_research_studies, " research studies conducted"),
               paste0(total_funded_conferences, " conferences hosted"),
               paste0(total_publications, " publications developed"))
colors_list <- c(color2, color3, color4, color5)

df <- create_callout_df(icons = icons_list, text = text_list, colors = colors_list)


create_callout_chart(df)





```

